What can Issues on GitHub and Google Code tell us about Bug Hazards in Android and Java?

### 2015-02-12 android-stacks.tex
changes on abstract.

This study performs a post mortem analysis of the exception stack traces 
included in issues
reported on Android projects hosted in GitHub and Google Code. 
The goal of this study is to investigate whether the reported exception stack traces
 can reveal common fragilities in the exception-related code.  
We call ``fragilities'' characteristics of the exception-related code 
that increase the likelihood of introducing failures such as uncaught 
exceptions -  leading to a decrease in application robustness.

@@

Hence, the study results are relevant to Android and 
Java developers who may underestimate how 
 exceptions are related to programming mistakes, and how 
 tricky it can be to prevent the uncaught exceptions from third-party code
 in the absence of ``exception interfaces". Moreover, the results are relevant to 
designers of languages and tools to consider ways of reducing such fragilities 
in Android and Java environments.

@@

The goal of this exploratory mining study was to investigate whether the exception stack traces 
embedded on issues of Android projects (hosted in GitHub and Google Code), can reveal 
\emph{common fragilities} on the exception code of both the applications and the underlying  framework. 
Our study focused on open-source apps, since the information needed to perform our study
cannot be retrieved from commercial apps, whose issue report systems and 
source codes are not publicly available. The Android open-source apps were also 
target of other research studies ~\cite{Linar13,ahimed}.

This exploration is guided by the set of best practices compiled in this study; we investigate whether characteristics 
of exception stack traces and the related source code it can pinpoint best practices violations. As mentioned before, 
we call  ``fragilities" characteristics on the exception-related code that favor the introduction
of failures such as uncaught exceptions and unintended handling.

To support this investigation, we developed a tool called ExceptionMiner (Section~\ref{sec:exceptionminer})
which extracts the exception stack traces embedded on issues, 
and combines stack trace information with source code and bytecode 
analysis. Such combination was needed to enable the identification of 
fragilities on the exception-related code. Moreover, manual inspections were also used to leverage
 the understanding of stack traces and support further discussions and insights.
In this study we explore the domain quantitatively and highlight interesting cases by 
exploring cases qualitatively.  Figure~ref{fig:overview} gives an overview of the study, whose tasks are detailed next.

@@

Based on the assumption that regardless the issue type, every exception stack
trace contains relevant information concerning the exception structure of the
projects analyzed - and therefore can review fragilities on the exception-related code -  
we opted for not restricting the analysis only on defect
issues\footnote{We conducted the same analysis on the defect issues and the top
exceptions are similar to the ones mentioned on defect issues. Due to space
limitation we limit to present the general analysis here. More detailed analysis
on the defect issues can be found at:
\url{https://github.com/souzacoelho/exceptionminer}}.
 Moreover, issues and pull requests are dual on Github; all pull requests have a corresponding 
``backing'' issue (i.e., an issue that is automatically generated per pull request). Therefore, we excluded such automatically generated
issues from our analysis.

@@

This section presents the results of the study, providing both a
quantitative and qualitative analysis of the outcomes. The structure of this section
 is centered on questions that guided the search for fragilities on the exception-related code
based on exception stack trace information.

@@

\emph{\textbf{On the Presence of Checked Exceptions on Exception Interfaces.}} 
In this study, the exception stack trace analysis revealed an unexpected fragility: a 
checked exception thrown by a native method and not declared on the exception 
interface of these methods signaling them. Such native method was defined on
 the Android platform -  which uses Java Native Invocation (JNI) to access 
native C/C++ code. This exception was thrown by the method getDeclaredMethods defined in java.lang.Class. 
The Java-side declaration of this method does not have any throws clause, 
leading programmers and the compiler to think that no checked exceptions can be thrown.
 However, the C-code implementation did throw a ``checked exception" called NoSuchMethodException, 
violating the declaration. The Java compiler could not detect this violation, because it does 
not perform static exception checking on native methods. This type of bug is hard to diagnose
because the developer usually do not have access to the native implementations. 
Consequently, since it is not expected by the programmer, when such method throws 
this exception, such undocumented exception may remain
uncaught and cause the app to crash, or may be mistakenly handled by subsumption.
The exception stack traces reporting this scenario is actually a real bug of Android 
Gingerbread version (which still accounts for 13.6\% of devices running Android).
%Among the Android projects on which this exception stack trace was reported we can cite:
%PageTurner (206 stars), ActionBarSherlock (6,041 stars), otto (1,618 stars).

@@

\emph{External Validity.} Our work uses GHTorrent dataset; although 
comprehensive and extensive is not an exact replica of Github. 
However, the result of this study does not depent on the analysis of
a complete Github dataset. Instead, the goal of our study was to 
pinpoint fragilities on the exception-related code based on 
exception stack trace mining of a subset of projects.
We limited our analysis to a subset of existing open-source Android projects.
We are aware that the exception stack traces reported 
for commercial apps can be different from the ones found in this study, and that
this subset is a small percentage of existing apps.
Such threats are similar to the ones imposed to other empirical studies 
which also used free or open-source Android apps ~\cite{Linar13,McDon13,ahimed}.
Moreover, several exception stack traces that support the findings of this study
refered to exceptions coming from methods defined on Android Application Framework
and third-party libraries.  Additionaly,  the ``fragilities" observed in this study are due to
characteristics of Java exception model, which can impose challenges 
the robustness of not only to Android apps but also to other systems
 based on the same exception model.

@@

\textit{Empirical Studies on Exception Handling Defects.} 
Cabral and Marques~\cite{cabral2007exception} analyzed the
source-code of 32 open-source systems, both for Java and .NET. They
observed that the actions inside handlers were very simple (e.g., logging and present a
message to the user). Coelho et al.  ~\cite{coelho2008assessing} performed an 
empirical study considering the fault-proneness of aspect-oriented implementations 
for handling exceptions. Two releases of both Java and AspectJ implementation of a 
product line were assessed as part of that study. Based on the use of an exception
flow analysis tool, the study revealed that the AOP  refactoring increased the 
number of uncaught exceptions, degrading the robustness of the AO version of every analyzed system.
The main limitation of approaches based static analysis based approaches are the number of false
positives they can generate, and the problems the faced when dealing with reflection libraries 
and dynamic class loading. Pingyu and Elbaum~\cite{Zhang12} were the first to perform
an empirical investigation of issues, related to exception-related bugs, on Android projects.  
They perform a small scale study on which they manually inspected the issues of 
5 Android applications. They observed that 29\% had to do with poor
exceptional handling code, this empirical study was used to motivate the development of a tool
aiming at amplifying existing tests to validate exception 
handling code associated with external resources. This work inspired ours,
 which automaticaly mined the exception stack traces embedded on issues 
reported on 639 open source Android projects. The goal of our study was
to identify common fragilities on the exception related code that can lead to 
failures such as uncaught exceptions.

@@

In this paper we present an exploratory study in which we mined the stack 
traces embedded in all issues defined in 482 Android projects hosted in Github and 
157 projects hosted in Google Code. Overall it included 6,005 exception stack traces.
In this study, the information extracted from stack traces was used in combination 
with source code and bytecode analysis to identify common characteristics of exception 
stack traces to pinpoint fragilities on the exception-related code. 
We call ``fragilities" characteristics on the exception-related code that favor the introduction
of  uncaught exceptions and unintended handling - which can contribute to 
degrade the robustness of applications.
Some fragilities consistently detected in this study were: 
(i) Approximately 52\% of the reported stack traces can be attributed to programming mistakes. 
In particular, 27.71\% of all stack traces reported a java.lang.NullPointerException as its root cause;
(ii) 267 stack traces were caused by undocumented runtime exceptions thrown by third party code caused;
(iii) undocumented checked exceptions thrown by JNI interface,
(iv) unexpected exception wrappings (e.g., Errors being wrapped in checked exceptions) 
revealing that Java hybrid exception model is not fully used according to its purpose.
Such fragilities can negatively affect not only the robustness of Android application 
but also the robustness of any Java-based application. 
Our results calls the attention for the need of tools and language support to help 
developers when dealing with such fragilities.

>>>>

This study performs a post mortem analysis of the exception stack traces 
included in issues
reported on Android projects hosted in GitHub and Google Code. 
The goal of this study is to investigate whether the reported exception stack traces
 can reveal common bug hazards in the exception-related code.  
A bug hazard  ~\cite{binder2000testing}  is a circumstance that increases the 
chance of a bug to be present in the software. An example of a bug hazard can 
be a characteristic of  the exception-related code which can increase the likelihood 
of introducing uncaught exceptions (one of the main causes of application crashes).

@@

Hence, the study results are relevant to Android and 
Java developers who may underestimate how 
 exceptions are related to programming mistakes, and how 
 tricky it can be to prevent the uncaught exceptions from third-party code
 in the absence of ``exception interfaces". Moreover, the results are relevant to 
designers of languages and tools to consider ways of dealing with such bug hazards 
in Android and Java environments.

@@

The goal of this exploratory mining study was to investigate whether the exception stack traces 
embedded on issues of Android projects (hosted in GitHub and Google Code), can reveal 
\emph{bug hazards} on the exception-related code of both the applications and the underlying framework. 
This exploration is guided by the set of best practices compiled in this study; we investigate whether characteristics 
of exception stack traces and the related source code it can pinpoint best practices violations and therefore to bug hazards.
As mentioned before, in this context ``bug hazards" are the characteristics on the exception-related code 
that favor the introduction of failures such as uncaught exceptions.

To support this investigation, we developed a tool called ExceptionMiner (Section~\ref{sec:exceptionminer})
which extracts the exception stack traces embedded on issues, 
and combines stack trace information with source code and bytecode 
analysis. Moreover, manual inspections were also used to leverage
 the understanding of stack traces and support further discussions and insights.

Our study focused on open-source apps, since the information needed to perform our study
cannot be retrieved from commercial apps, whose issue report systems and 
source codes are not publicly available. The Android open-source apps were also 
target of other research studies ~\cite{Linar13,ahimed}.

In this study we explore the domain quantitatively and highlight interesting cases by 
exploring cases qualitatively.  Figure~ref{fig:overview} gives an overview of the study, whose tasks are detailed next.

@@

Based on the assumption that regardless the issue type, every exception stack
trace contains relevant information concerning the exception structure of the
projects analyzed - and therefore can review bug hazards on the exception-related code -  
we opted for not restricting the analysis only on defect
issues\footnote{We conducted the same analysis on the defect issues and the top
exceptions are similar to the ones mentioned on defect issues. Due to space
limitation we limit to present the general analysis here. More detailed analysis
on the defect issues can be found at:
\url{https://github.com/souzacoelho/exceptionminer}}.
 Moreover, issues and pull requests are dual on Github; all pull requests have a corresponding 
``backing'' issue (i.e., an issue that is automatically generated per pull request). Therefore, we excluded such automatically generated
issues from our analysis.

@@

This section presents the results of the study, providing both a
quantitative and qualitative analysis of the outcomes. The structure of this section
 is centered on questions that guided the search for bug hazards on the exception-related code
based on exception stack trace information.

@@

\emph{\textbf{On the Presence of Checked Exceptions on Exception Interfaces.}} 
In this study, the exception stack trace analysis revealed an unexpected bug hazard: a 
checked exception thrown by a native method and not declared on the exception 
interface of these methods signaling them. Such native method was defined on
 the Android platform -  which uses Java Native Invocation (JNI) to access 
native C/C++ code. This exception was thrown by the method getDeclaredMethods defined in java.lang.Class. 
The Java-side declaration of this method does not have any throws clause, 
leading programmers and the compiler to think that no checked exceptions can be thrown.
 However, the C-code implementation did throw a ``checked exception" called NoSuchMethodException, 
violating the declaration. The Java compiler could not detect this violation, because it does 
not perform static exception checking on native methods. This type of bug is hard to diagnose
because the developer usually do not have access to the native implementations. 
Consequently, since it is not expected by the programmer, when such method throws 
this exception, such undocumented exception may remain
uncaught and cause the app to crash, or may be mistakenly handled by subsumption.
The exception stack traces reporting this scenario is actually a real bug of Android 
Gingerbread version (which still accounts for 13.6\% of devices running Android).
%Among the Android projects on which this exception stack trace was reported we can cite:
%PageTurner (206 stars), ActionBarSherlock (6,041 stars), otto (1,618 stars).

@@

\emph{External Validity.} Our work uses GHTorrent dataset; although 
comprehensive and extensive is not an exact replica of Github. 
However, the result of this study does not depent on the analysis of
a complete Github dataset. Instead, the goal of our study was to 
pinpoint bug hazards on the exception-related code based on 
exception stack trace mining of a subset of projects.
We limited our analysis to a subset of existing open-source Android projects.
We are aware that the exception stack traces reported 
for commercial apps can be different from the ones found in this study, and that
this subset is a small percentage of existing apps.
Such threats are similar to the ones imposed to other empirical studies 
which also used free or open-source Android apps ~\cite{Linar13,McDon13,ahimed}.
Moreover, several exception stack traces that support the findings of this study
refered to exceptions coming from methods defined on Android Application Framework
and third-party libraries.  Additionaly,  the ``bug hazards" observed in this study are due to
characteristics of Java exception model, which can impose challenges 
the robustness of not only to Android apps but also to other systems
 based on the same exception model.

@@

\textit{Empirical Studies on Exception Handling Defects.} 
Cabral and Marques~\cite{cabral2007exception} analyzed the
source-code of 32 open-source systems, both for Java and .NET. They
observed that the actions inside handlers were very simple (e.g., logging and present a
message to the user). Coelho et al.  ~\cite{coelho2008assessing} performed an 
empirical study considering the fault-proneness of aspect-oriented implementations 
for handling exceptions. Two releases of both Java and AspectJ implementation of a 
product line were assessed as part of that study. Based on the use of an exception
flow analysis tool, the study revealed that the AOP  refactoring increased the 
number of uncaught exceptions, degrading the robustness of the AO version of every analyzed system.
The main limitation of approaches based static analysis based approaches are the number of false
positives they can generate, and the problems the faced when dealing with reflection libraries 
and dynamic class loading. Pingyu and Elbaum~\cite{Zhang12} were the first to perform
an empirical investigation of issues, related to exception-related bugs, on Android projects.  
They perform a small scale study on which they manually inspected the issues of 
5 Android applications. They observed that 29\% had to do with poor
exceptional handling code, this empirical study was used to motivate the development of a tool
aiming at amplifying existing tests to validate exception 
handling code associated with external resources. This work inspired ours,
 which automaticaly mined the exception stack traces embedded on issues 
reported on 639 open source Android projects. The goal of our study was
to identify common bug hazards on the exception related code that can lead to 
failures such as uncaught exceptions.

@@

In this paper we present an exploratory study in which we mined the stack 
traces embedded in all issues defined in 482 Android projects hosted in Github and 
157 projects hosted in Google Code. Overall it included 6,005 exception stack traces.
In this study, the information extracted from stack traces was used in combination 
with source code and bytecode analysis to identify common characteristics of exception 
stack traces to pinpoint bug hazards on the exception-related code. 
In this study, bug hazards are characteristics on the exception-related code that favor the introduction
of  uncaught exceptions and unintended handling - which can contribute to 
degrade the robustness of applications.
Some bug hazards consistently detected in this study were: 
(i) unexpected exception wrappings (e.g., Errors being wrapped in checked exceptions)  - 
revealing that Java hybrid exception model is not fully used according to its purpose;
(ii) 267 stack traces were caused by undocumented runtime exceptions thrown by third party libraries or the Android Platform;
and (iii) undocumented checked exceptions thrown by the Android Platform. 
Moreover, we could also observe that approximately 52\% of the reported stack traces 
can be attributed to programming mistakes; in particular, 27.71\% of all stack traces 
reported a java.lang.NullPointerException as its root cause;
The bug hazards discussed in this work can negatively affect not only the robustness of Android application 
but also the robustness of any Java-based application. 
Our results calls the attention for the need of tools and language support to help 
developers when dealing with them.
